### 2019-03-26 0_abstract.tex
Apply Alex's comments to abstract and introduction.

\begin{abstract}
% 149 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelops of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench consists of a configurable \emph{task graph} with one task per
point in an iteration space, and dependencies between tasks determined
by a \emph{dependence relation}. Tasks execute a variety of kernels,
including compute- and memory-bound, and dependencies carry payloads
of configurable size, enabling communication-bound scenarios. Multiple
task graphs can be executed to enable task parallelism.

We conduct a comprehensive study with implementations of Task Bench in
13 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of varying dependence patterns and numbers of
dependencies. We also study each system's scalability, and ability to
hide communication and mitigate load imbalance.
\end{abstract}

>>>>

\begin{abstract}
% 129 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench seeks to provide the ease of implementation and portability of
microbenchmarks, which makes comparisons between many systems
possible, with the ability to easily model a wide range of application
scenarios, giving it coverage approximating large benchmarks.

We conduct a comprehensive study with implementations of Task Bench in
13 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of scenarious with varying dependence patterns. We also
study each system's scalability, and ability to hide communication and
mitigate load imbalance.
\end{abstract}

### 2019-03-26 1_introduction.tex
Apply Alex's comments to abstract and introduction.

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models, languages and runtime systems. These systems often provide
capabilities such as asynchronous execution of units of work, called
\emph{tasks}, coordinated in some manner with data movement and
synchronization. But while the features of these systems are widely
written about in the literature, a large and often difficult to answer
question is how different systems will perform for a given
application.

One approach to answering this question has focused around the
development of \emph{proxy-} or \emph{mini-apps}. So called because
they embody the logic of a larger application, but in reduced form,
mini-apps are popular because they can provide insight without
incurring the costly expense of developing a full-blown application
code. However, despite the name, mini-apps still requite significant
investment to develop, especially to a level of quality need to reach
a publishable state. In many cases, the effort to tune for performance
exceeds the effort to develop the mini-app in the first place. This
can be exacerbated if the system in question is undergoing concurrent
development. As a result, implementations of mini-apps, though
well-intentioned, often do not reach a level of maturity required to
perform a broad comparison between systems, and as a result there are
very few comparative studies involving multiple programming systems
that reach publication.

A compounding issue is one of methodology: how does one predict how
the performance observed for a mini-app will generalize to a full
application, even one closely related? Part of the problem here is
that weak and strong scaling, which are by far the most commonly
reported metrics for performance of both mini- and full applications,
do nothing to elucidate the performance envelop of a particular
system. Weak scaling is notorious for being able to hide arbitrary
amounts of overhead, as long as the problem size is sufficiently
large. And strong scaling runs the risk of entangling unrelated
effects: namely, the baseline overhead of a given system, with the
growth of communication latencies, and potentially the saturation of
global resources such as communication and I/O bandwidth, when using
progressively larger portions of the machine.

To address the first of these issues, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of programming systems under a variety of conditions. A
critical property of Task Bench is that, while powerful, it is simple
enough to be implemented in a wide variety of systems with reasonable
effort, to a level of quality necessary for comparative study.

Task Bench consists of a configurable \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a broad space of application
performance behaviors.

To resolve the methodological issue, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively, METG is
the smallest task granularity which can be executed while maintaining
a given level of computational efficiency (e.g., as measured by the
percentage of peak FLOPS achieved). METG captures the essence of what
is important in a weak or strong scaling study: that is, the behavior
at the limit of scalability. For weak scaling, METG corresponds
directly to the smallest problem size that can be weak scaled with a
given level of efficiency. For strong scaling, METG can be used to
compute the point at which strong scaling can be expected to tail
off. In either case, METG is a useful, intuitive measure as it
encapsulates the task granularity at which systems can be relied on to
execute efficiently. As a rule of thumb, as long as an application's
average task granularity exceeds the METG, the application should
execute efficiently.

Together, METG and Task Bench help to mitigate issues which are common
in limit studies on the task scheduling throughput of runtime
systems. Such studies typically employ a metric called \emph{tasks per
  second} (TPS), which is almost universally measured with trivial
dependencies and empty tasks. While TPS is certainly an upper bound on
task scheduling throughput, it is not a tight bound, as the cost of
non-trivial dependencies can be significant and sometimes adversely
impact scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

But there is a deeper issue with the use of empty tasks. In all
realistic scenarios, runtime systems must share resources with the
application, either explicitly by e.g., allocating cores for runtime
use, or implicitly by sharing a time slice on the same cores used for
application tasks, and other resources such as memory bandwidth and
processor cache capacity. The use of empty tasks makes the application
share of these resources appear to be zero, which permits usage by the
runtime system to be unconstrained. How then is anyone supposed to
determine what resource usage to consider reasonable on the part of
the runtime system? Only by requiring the application usage of
resources to meet a certain minimum level, as in METG, can we ensure
that runtime system resource usage is constrained to realistic levels.

We conduct a comprehensive comparative study of Task Bench
implementations in 13 different programming systems, including
Chapel~\cite{Chapel07}, Charm++~\cite{Charmpp93}, MPI~\cite{MPI},
OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40},
PaRSEC~\cite{PARSEC13}, Realm~\cite{Realm14}, Regent~\cite{Regent15},
Spark~\cite{Spark10}, StarPU~\cite{StarPU11},
Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
X10~\cite{X1005}. Using the METG metric, we determine the baseline
runtime overhead of each system along with overheads associated with
increasing number and/or variations in the pattern of dependencies. We
also explore each system's ability to scale, hide communication
overhead, and mitigate load imbalance.

@@

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 13 programming systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench performance on up to 256 Haswell nodes of the Cori
  supercomputer.
\end{itemize}

>>>>

The challenge of parallel and distributed computation has lead to a
wide variety of proposals for programming models, languages, and
runtime systems. Some, like MPI~\cite{MPI}, have achieved widespread
usage while many more are in the research and development stage. Our
goal in this paper is to develop a meaningful and useful framework for
comparing the performance of parallel and distributed programming
systems.

% note: removed definition of task

One approach to comparing the performance of different systems has
focused on the development of \emph{proxy-} or
\emph{mini-apps}. Because they embody the logic of a larger
application, but in reduced form, mini-apps can provide insight
without incurring the costly expense of developing a full-blown
application code. However, despite the name, our experience is that
mini-apps still requite significant investment to develop, especially
to a level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop the
mini-app in the first place. As a result, implementations of mini-apps
often do not reach a level of maturity required to compare
systems. Indeed, there are few such published studies, and none to our
knowledge that involve more than a handful of systems.

There is another important issue: How does one predict how
the performance of a mini-app generalizes to a full
application, even one closely related? Part of the problem here is
that weak and strong scaling, which are by far the most commonly
reported metrics for performance of both mini- and full applications,
do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead with sufficiently large problem sizes. And strong
scaling does not separate the overhead of the underlying runtime
system from the growth in communication latencies, and potential
saturation of resources such as communication and I/O bandwidth, when
using progressively larger portions of the machine.

To address the first of these issues, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative studies.

A key insight of Task Bench is that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication required for
distributed execution. Task Bench consists of a configurable \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a large space of application
behaviors.

To isolate the contribution of the programming model's runtime system
to application performance, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively,
METG(50\%) is the smallest application task granularity which can be
executed while maintaining at least 50\% computational efficiency,
meaning that the application achieves at least 50\% of peak
performance. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem
size that can be weak scaled with 50\% efficiency. For strong scaling,
METG(50\%) can be used to compute the scale at which performance can
be expected to dip below 50\% efficiency. A good rule of thumb is that
as long as an application's \emph{average} task granularity exceeds
METG(50\%), the application should achieve at least 50\% efficiently.

Task Bench and METG mitigate issues common in limit studies of runtime
systems for parallel and distributed programming. Such studies often
employ the metric \emph{tasks per second} (TPS), which is almost
universally measured with trivial (i.e., no) dependencies. While
phrased in terms of tasks, TPS can be measured for any system as long
as the application in question has identifiable units of work that run
to completion without interruption. TPS is an upper bound on
runtime-limited application throughput, but not a tight bound, as the
cost of non-trivial dependencies can be significant and may impact
scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

There is another, deeper issue with TPS. TPS may be measured with
empty tasks~\cite{LegionTracing18} or with tasks of some
duration~\cite{Canary16, Armstrong14}. If empty tasks are used, the
resulting upper bound on task scheduling throughput fails to represent
any realistic application, as no
useful work is being performed. If non-empty tasks are used, \emph{but
  the efficiency of the overall application is not reported}, then TPS
is not a measurement of runtime-limited performance. Large tasks may
be used to hide any amount of runtime overhead,
while tasks that are too small may result in a drop in total
application throughput even as TPS increases. Only by constraining
efficiency, as in METG, can we measure runtime-limited application
throughput while performing useful application work.

We conduct a comprehensive comparative study of Task Bench
implementations in 13 different programming systems, including
Chapel~\cite{Chapel07}, Charm++~\cite{Charmpp93}, MPI~\cite{MPI},
OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40},
PaRSEC~\cite{PARSEC13}, Realm~\cite{Realm14}, Regent~\cite{Regent15},
Spark~\cite{Spark10}, StarPU~\cite{StarPU11},
Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
X10~\cite{X1005}. Using METG(50\%), we determine the baseline
runtime overhead of each system along with overheads associated with
increasing numbers of and/or variations in the pattern of dependencies. We
also explore each system's ability to scale, hide communication
overhead, and mitigate load imbalance.

@@

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 13 programming systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench performance on up to 256 Haswell nodes of the Cori
  supercomputer.
\end{itemize}
